# Reproducible Research: Peer Assessment 1
First of all, we load the required libraries:

```{r}
library(lattice)
```



## Loading and preprocessing the data
We read the data and we cast the dates:

```{r}
data <- read.csv("activity.csv", stringsAsFactors = FALSE)
data$date <- as.Date(data$date, "%Y-%m-%d")
```


## What is mean total number of steps taken per day?

We calculate the sumation of the steps per day as follows (we consider the summation of NAs as 0):

```{r}
dataSumPerDay<- aggregate(data$steps, list(Date = data$date), sum, na.rm = T)
colnames(dataSumPerDay)[2] <- "summation"
```

In the following figure we can see the histogram of the resultant data:
```{r fig.width=8, fig.height=6}
hist(dataSumPerDay$summation, main ="Steps per day", xlab ="sum(Steps)")
````

Next, we calculate the mean and the median of the summation: 
```{r}
meanSummation <- mean(dataSumPerDay$summation)
medianSummation <- median(dataSumPerDay$summation)
meanSummation
medianSummation
```

## What is the average daily activity pattern?

We calculate the daily step average pattern as follows:  

```{r}
dataMeanPerInterval <- aggregate(data$steps, list(Date = data$interval),mean,na.rm = T)
colnames(dataMeanPerInterval) <- c("interval", "meanStepPerIntervals")
```

We can see the plotting in the next figure:
```{r fig.width=8, fig.height=6}
plot(dataMeanPerInterval$interval, dataMeanPerInterval$meanStepPerIntervals, type = "l", xlab = "Interval", ylab = "Step Mean", main = "Step average by interval")
```

The interval with the maximum number of steps is:
```{r}
posMax <- match(max(dataMeanPerInterval$meanStepPerIntervals), dataMeanPerInterval$meanStepPerIntervals)
intervalMaxSteps <- dataMeanPerInterval$interval[posMax]
intervalMaxSteps
```

## Imputing missing values
We will fill the missing values(NA) with the mean for that 5-minute interval.  
First we count the number of NAs:

```{r}
sum(is.na(data$steps))
```

Then we replace the NAs with the means and we copy it to a new data frame:

```{r}
tmp <- mapply(function(x, y) {
    if(is.na(x))
        dataMeanPerInterval$meanStepPerIntervals[match(y, dataMeanPerInterval$interval)]
    else
        x
    }, data$steps, data$interval)

dataNoNA <- data
dataNoNA$steps <- tmp
```

We recalculate the step summation for each day and we plot it as we do in the first section. We can see the difference between the two histograms:

```{r fig.width=8, fig.height=6}
dataNoNASumPerDay<- aggregate(dataNoNA$steps, list(Date = dataNoNA$date), sum, na.rm = T)
colnames(dataNoNASumPerDay)[2] <- "summation"
hist(dataNoNASumPerDay$summation, main = "Steps per day", xlab ="sum(Steps)")
```

## Are there differences in activity patterns between weekdays and weekends?

Finally we are going to see the difference between the weekend and the weekday patterns.

First, we create a factor that splits the data in two subsets, weekend and weekdays. (I haven't used the weekdays() function because the configured language of my R studio is in Spanish):

```{r}
listTmp <- c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", 
  "Friday", "Saturday")[as.POSIXlt(dataNoNA$date)$wday + 1]

dataNoNAWeeks <- cbind(dataNoNA,listTmp)
colnames(dataNoNAWeeks)[4] <- "weekday"

factorWeek <- as.factor(dataNoNAWeeks$weekday %in% c("Sunday","Saturday"))
levels(factorWeek) <- list( Weekends = TRUE, Weekdays= FALSE)

dataNoNA <- cbind(dataNoNA, factorWeek)
```

Now, we calculate the interval average of both subsets:

```{r}
dataMeanIntervWeekday <- aggregate(dataNoNA$steps, list(Date = dataNoNA$interval, dataNoNA$factorWeek), mean, na.rm = T)
colnames(dataMeanIntervWeekday) <- c("interval", "weekday", "averageSteps")

```

We can see the time series plot in the next figure:
```{r fig.width=8, fig.height=6}
xyplot(averageSteps~interval|weekday, main="Difference Between Patterns", ylab="Number of steps", xlab="Interval", data = dataMeanIntervWeekday, type = 'l', layout = c(1,2))
```
